% !TEX root = ../main.tex

\begin{acknowledgements}
When I received the offer letter from SJTU, my thrill for the upcoming study life was accompanied by underneath disquiet. By that time, as an undergraduate student with little experience in hardware, coding, and academic writing, I was not able to picture myself as a senior Ph.D. candidate who can lead research projects, comprehend paper writing, and give junior students advice. Learning of these skills and my mental improvements are inseparable from the people who genuinely helped me during the past five years:

My acceptance by Prof. Peter Bradley Shull was a true honor. I received his strong support and generous help when I realized and transferred my research interest to data-driven gait kinetics, when I decided to pursue a Ph.D. degree in the middle stage of my master program, and when I made up my mind to start a career in academia in the late stage of my Ph.D. program. Also, his exceptional and consistent pursuit of journal paper quality deeply influenced me, and I will try to keep it this way for my future publications.
% Try to be helpful in dealing affairs from daily laboratory affairs to international collaborations 
% so that he only needs to observe but not participate
% from the perspective of xx expert.

Dr. Eni Halilaj's expertise in machine learning has notably promoted our last knee moment estimation project. Participating in her research group's weekly online meeting was such a special experience that gave me a chance to be updated with the latest advances in the biomechanical field. 
% Also, xx from communication with her brilliant students who have a strong computer science background.
% Musculoskeletal Biomechanics Lab, Carnegie Mellon University. 


Prof. Xiangyang Zhu, Xinjun Sheng, and Wei Dong set examples of being distinguished professors. They provided adequate financial support, gave constructive advice, and offered a free research environment. I feel indebted to them for sharing such a well-equipped laboratory. Dr. Hai Hu, an expert in orthopedics, inspired me to use my work to help patients.

My parents always try to offer me generous financial support, which I declined most of the time. They also taught me to have big-picture thinking and envision for the long term, and at the same time, they highly respect my decisions.

Yangjian Huang positively affected me with his spirit of pragmatism and sense of humor. Junkai Xu corrected many confusing mistakes of my research with his rich knowledge and strong will. Haisheng Xia provided irreplaceable help in my resonant skin stretch project using his extraordinary ability in dealing with hardware problems. Shuo Jiang set a hardworking and example for our research team. Bingfei Fan inspired me with his academic thinking. Xinyu Song gave me helpful company both emotionally and academically. Fangyuan taught me the operation of Vicon, which is a useful technique I have passed to more than ten other students. Zachary Strout patiently collaborated with me on the Huawei project, which I cannot finish alone. Dianxin Wang disciplined me to code more concisely, and I disciplined him to write more formally. David Chiasson gave me many helpful grammar suggestions for my first Journal paper.

Thank Zhiyong Wang, Heng Zhang, Sensen Liu, Huanpen Ye, Gang Chen, Wei Xia, Andong Yi, Meng Wang, Daiyan Pen, Qinghua Gao, Peiqi Kang, Jinxuan Li, Yuan Lin, Jue Ouyang, Changyang Gong, Hongzhe Yu, Sen Xu, Yansong Wang, Yuyang Zhao, Yukun Wang, Xing Zhang, and Ruoyu Yang for their companies in my five year Ph.D. journey.

Thank Haifeng Ding, Xining Lv, Lin Ji, Qihao Lu, Qiyuan Lu for many memorable drinking and bragging nights.

\end{acknowledgements}
