

\chapter{Publications}

\large
\noindent \textbf{Journal Articles}
\normalsize

\begin{hangparas}{.25in}{1}
[1] 
\textbf{Tian Tan}, Zachary A. Strout, and Pete B. Shull, \href{https://ieeexplore.ieee.org/document/9162450}{
``Accurate Impact Loading Rate Estimation During Running via a Subject-Independent Convolutional Neural Network Model and Optimal IMU Placement,''} \textit{\textbf{IEEE Journal of Biomedical and Health Informatics}}, vol. 25, pp. 1215-1222, 2021. (SCI, IF=5.2)

[2] 
\textbf{Tian Tan}, Zachary A. Strout, Haisheng Xia, Mostafa Orban, and Pete B. Shull, 
\href{https://ieeexplore.ieee.org/document/9309100}{``Magnetometer-Free, IMU-Based Foot Progression Angle Estimation for Real-Life Walking Conditions,''} \textit{\textbf{IEEE Transactions on Neural Systems \& Rehabilitation Engineering}}, vol. 29, pp. 282-289, 2021. (SCI, IF=3.3)

[3]
\textbf{Tian Tan}, David P. Chiasson, Hai Hu, and Pete B. Shull, \href{https://www.sciencedirect.com/science/article/pii/S0021929019306633}{
``Influence of IMU Position and Orientation Placement Errors on Ground Reaction Force Estimation,''} \textit{\textbf{Journal of Biomechanics}}, vol. 97, p. 109416, 2019. (SCI, IF=2.3)

[4] 
Pete B. Shull, \textbf{Tian Tan}, Heather M. Culbertson, Xiangyang Zhu, and Allison M. Okamura, \href{https://ieeexplore.ieee.org/document/8716295}{``Resonant Frequency Skin Stretch for Wearable Haptics,''} \textit{\textbf{IEEE Transactions on Haptics}}, vol. 12, pp. 247-256, 2019. (SCI, IF=3.1)

[5]
王钿鑫,  \textbf{谈天}, Pete B. Shull, \href{http://www.mechanobiology.cn/yyswlx/article/abstract/20210122001}{
``基于惯性传感网络的穿戴式步行膝关节力矩估计.''} \textit{\textbf{医用生物力学}}, in press, 2021.


\vspace{1cm}

\end{hangparas}

\large
\noindent \textbf{Conference Abstracts}
\normalsize

\begin{hangparas}{.25in}{1}
[1] 
\textbf{Tian Tan}, Dianxin Wang, Pete B. Shull, Eni Halilaj, \href{https://drive.google.com/file/d/1RI4F07uvpwLE_gAOshSKr4ogJfCXAtHS/view?usp=sharing}{
``IMU-Camera Fusion for Estimating Knee Adduction and Flexion Moments,''} American Society of Biomechanics (ASB) Conference, 2021.

[2] 
\textbf{Tian Tan}, Zachary A. Strout, Pete B. Shull, 
\href{https://drive.google.com/file/d/1u_1mprioTqSyj5kP6kDDRL62R3FbT_DD/view?usp=sharing}{
``Magnetometer-free inertial-sensor-based foot progression angle estimation,''} American Society of Biomechanics (ASB) Conference, 2020.

[3] Yangjian Huang, Xinyu Song, \textbf{Tian Tan}, Pete B. Shull, ``Motion tracking based on embedded visual-inertial coupled SLAM for upper-limb movement training,'' IEEE International Conference on Wearable and Implantable Body Sensor Networks (BSN), 2018. 

\vspace{1cm}

\end{hangparas}

\large
\noindent \textbf{Under Review Papers}
\normalsize

\begin{hangparas}{.25in}{1}
[1] \textbf{Tian Tan}, Dianxin Wang, Pete B. Shull, Eni Halilaj, ``IMU and Smartphone Camera Fusion for Knee Adduction and Knee Flexion Moment Estimation During Walking,'' \textit{\textbf{IEEE Transactions on Industrial Informatics}}, under review

\end{hangparas} 
